Study of Intra- and Inter-Job Interference with Different Job allocations

### 2015-10-14 submission.tex
minor change

Network contention between concurrently running jobs on HPC system is one of the primary causes of performance variability. Optimizing job allocation and avoiding network sharing are proven techniques to alleviate network contention and performance degradation. In this work, we propose a novel job communication pattern aware allocation strategy which serves as a critical module in our future HPC batch scheduler. The novelty of our new strategy is that it makes allocation decisions based on job’s communication pattern in addition to system topology and resource availability information. The scheduler uses job communication patterns information to decide whether its performance relies heavily on the network or not, and what the potential inter/intra-job interference would be for a given allocation. Thus, the scheduler can pick a preferable resource allocation based on a specific job's communication pattern. We validate the effectiveness of our new design using traces collected from three representative parallel applications. The idea presented in this paper is widely applicable: HPC parallel applications have common communication patterns, and allocating resource with communication pattern awareness can improve performance on these systems with a variety of network topologies.

@@

%\textcolor{red}{
%Partition-based allocation achieves good job performance by sacrificing system utilization, whereas non-contiguous allocation can result in better system utilization but could severely degrade performance of user jobs. What is needed is a more flexible strategy that provides allocations that minimize intra/inter-job contention but allow greater system utilization. In this work, we propose a guideline for future batch scheduling systems, a smart way to allocate resources with the awareness of jobs communication patterns.
%
%Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy. Whether it use partition-based or non-contiguous policy, means it chooses the benefit of either system or job over the other. Some propose compromised policies like buddy system\cite{buddy system paper} still make the allocation for job blindly without knowing job's communication information. 
%
%Although HPC applications complexity continue to grow, people could get more knowledge about the applications communication information with the help from various parallel application profiling tools, such as TAU\cite{tau}, MPIP\cite{mpip} and DUMPI\cite{sst}. Job's communication trace and performance analysis can be get by using such tool at runtime. 
%
%The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions. This is because the communication pattern of parallel applications on HPC system can be summarized into a couple of categories\cite{hpdc2015-ornl}. The deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern. Job's communication traces study would be extreme helpful to leadership computing facility too. The set of applications running on those leadership computing facilities is usually stable, and the workloads from those HPC systems is usually has great repetitiveness. We analyzed the workload on Mira, an IBM Blue/Gene Q system at Argonne Leadership Computing Facility, found that most jobs require 2k nodes and about 14 particular jobs consist of the majority its workload\ref{fig: repetitiveness of Mira}. In other word, Mira accommodates mostly some particular 14 jobs and most jobs have the same size.
%}

@@

%A parallel application usually conforms to a combination of several basic communication patterns\cite{hpdc2015-ornl}. At its different execution phases, the application's communication behavior may follow different basic patterns respectively. \textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution. Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc. In this work, we focus on the locality of application's rank-to-rank communication topology graph.

A parallel application usually conforms to a combination of several basic communication patterns \cite{hpdc2015-ornl}. At its different execution phases, the application's communication behavior may follow different basic patterns respectively. There are many profiling tools \cite{tau}\cite{mpip}\cite{scala}\cite{sst}\cite{oxbow} available to capture information regarding communication patterns of parallel applications. They can help analyze parallel applications' communication behavior, providing information like the percentage of different MPI operations of the applications, communication topology, the amount data transferred between processes.

@@

The second MiniApp we have is CrystalRouter, the extracted communication kernel of the full application Nek5000 \cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory. It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos. CrystalRouter demonstrates the "many-to-many" communication pattern through scalable multi-stage communication process. The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.

The collective communication in CrystalRouter utilizing a recursive doubling approach. Ranks in CrystalRouter conform to a n-dimensional hypercube and recursively splitting into (n-1)-dimensional hypercubes, with communication happening with respect to the links that are severed. The pattern of this recursive communication can be found in Figure \ref{fig: cr communication topology}. By nature of the logarithmic distribution, more data transferring will be conducted within the local neighborhood, which consists of 5 ranks.

@@

MultiGrid represents a group of applications whose dominant communication pattern is Many-to-Many (some literatures also refer it as All-to-All \cite{hpdc2015-ornl}).  Applications with such similar dominant communication pattern are like FillBoundary and MiniDFT \cite{design forward webpage}.

@@

%\textcolor{red}{Apps Summary} These three applications have representative communication patterns that are commonly seen in the HPC system workload. AMG represents a group of applications whose dominant communication pattern is Nearest Neighbor. MultiGrid represents applications whose dominant communication pattern is Many-to-Many (some literatures also refer it to All-to-All\cite{hpdc2015-ornl}). CrystalRouter represents those applications whose dominant communication is a hybrid of multi-stage local communication and hierarchical global communication.

@@

It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context. There are a couple of reasons for that. First, the allocation strategy used on production machine is part of the system software, which can not be changed by user. Even the system administrator is not authorized to make that change. Second, it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference then compare the results to that in the presence of interference. Due to those reasons, we resort to simulation for this work.

@@

A simulation toolkit named CODES (Co-Design of Multi-layer Exascale Storage Architecture) from Argonne National Lab enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}. CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}. CODES support both torus and dragonfly network with high fidelity flit-level simulation. CODES has this network workload component that capable of conducting trace-driven simulation. It can take real MPI application trace generated by SST DUMPI \cite{sst} to drive CODES network models. In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems. Since torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
%We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies. The new API can help us to explore the impact of different allocation strategies to those three applications in Section \ref{sec:application study}

@@

The topology of torus network is k-ary n-cube, with $k^n$ nodes in total arranged in an n-dimensional grid having k nodes in each dimension. Each node has 2$\times$ n direct linked neighbor nodes. The torus network performance is determined by its dimensionality and link bandwidth. As the increase of the dimensionality of torus network, so does the number of links connected with each node. The increased aggregated bandwidth of each node will definitely reduce the data transfer time of each rank in the application. Figure \ref{fig:dimensionality-study} shows the performance of the AMG, CrystalRouter and MultiGrid on a 2K node torus network model with a 3D torus (16x16x8), a 5D torus (8x4x4x4x4), and a 7D torus (4x4x4x4x2x2x2). The bandwidth of direct link between nodes is 2GiB/s in each direction \cite{bgq}, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus.

We can see from Figure \ref{fig:dimensionality-study} that the data transfer time of those three applications are greatly reduced as the dimensionality increases. The increased aggregated bandwidth of each node can accelerate the data transfer. We allocate each rank within each application linearly on the 3D torus. However, due to the variance of data transfer volume of each rank, AMG's data transfer time has great dispersion, shown as the outliers above the boxes in Figure \ref{fig:dimen-amg}. There is no such great dispersion of MultiGrid, shown in Figure \ref{fig:dimen-mg}, since there is no great variance of data transfer of each rank. The data transfer of each rank in MultiGrid varies slightly  between 2.8MB to 2.9MB. \textcolor{red}{explain crystalrouter}.

@@

Communication variability due to network sharing can cause application performance degradation in two ways, namely intra- and inter-job interference. The \textbf{\emph{intra-job interference}} refers to the network contention between ranks within each application. There are lots of research works \cite{jingjin}\cite{ozan} try to come up with better task mapping algorithms to alleviate the intra-job interference, which is out the scope of our work. In this work, we focus on the analysis of such interference with different allocations.

\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources. Such interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}. Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than partition-based system. Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to inter-job interference when running on non-contiguous allocation \cite{abhinav-sc13}.

In this section we will study both kinds of interference with with AMG, CrystalRouter and MultiGrid in their medium scale. To be specific, we choose AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks and simulate their running on 2K-node (16 $\times$ 16 $\times$ 8) 3D torus network.

We can get observations that 
\begin{enumerate}
    \item The compact allocation may not guarantee the best performance for every application. 
    \item The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefer compact allocation. 
    \item The allocation unit size should be decided according to application's communication pattern. Unit size that big enough to preserve the locality of the application will result better performance. 
\end{enumerate}

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/cont/allocation}
        \caption{ }
        \label{fig:cont_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/inter-job/noncont/sharing}
        \caption{ }
        \label{fig:noncont_sub1}
    \end{subfigure}%
    \caption{(a) Illustrates the contiguous allocation for 64-node job in three different shapes. ``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube. (b) Illustrates the non-contiguous allocation. Different colors represent different jobs. The nodes assigned to different jobs are interleaved, the allocation unit is 4-node.}
\end{figure}

@@

To study the intra-job interference of each job, we simulate each job running on 3D torus network with different allocation shapes. We assign each application with three different shapes allocation, i.e. 3D balanced-cube, 3D unbalanced-cube, 2D mesh.

@@

There are lots of research work try to design fancy allocation algorithms to provide application with most compact allocation. However, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application. Figure \ref{fig:shapstudy-amg} shows that for AMG, who has this symmetric local communication pattern through all ranks, spent less time when assigned with 3D-unbalanced allocation than 3D-balanced. And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}. Since MultiGrid's communication pattern is many-to-many dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message between ranks in MultiGrid. CrystalRouter has with both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.

@@

We simulate three application running concurrently with different allocation to analysis the interference between them. First, we assign each application with contiguous allocation side by side, and compare their performance with each application running exclusively. For AMG, sharing the network with other application won't affect its performance, since all the data transfer are between neighboring ranks. CrystalRouter and MultiGrid will suffer prolonged communication time when sharing network with others. The results are shown in Figure \ref{fig: concurrent-cont-time-box}.

Then, we assign each application with non-contiguous allocation and make them running concurrently. The allocation unit belongs to different jobs are interleaved as shown in Figure \ref{fig:noncont_sub1}. The unit size is critical to the application's performance. Figure \ref{fig: 3apps interjob study} shows the results of each application data transfer time with different allocation unit size.

@@

\section{Discussion}
\label{sec:discussion}
HPC system usually accommodate many parallel jobs running concurrently. Each job gets the required number of computing nodes by system resource manager. The allocated nodes for each job communicate with each other through system interconnect network in a sharing way, which cause interference as a negative side-effect. This interference can introduce performance degradation to both job and system, such as prolonged job running time and reduced system throughput.

@@

The most promising way to utilize such knowledge is to make smart allocation based on job's dominant communication pattern. Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy. Whether it uses partition-based or non-contiguous policy, it chooses the benefit of either system or job over the other. Some proposed compromised policies \cite{lo}\cite{pascual}\cite{yang-cluster14} to use fix allocation unit size and  make the allocation for job blindly without knowing job's communication information.

providing insights for ....
Based on the findings in this work, we propose the idea of smart job allocation with consideration about job's communication pattern. As the dominant communication pattern of parallel applications on HPC system can be summarized into a couple of categories, the deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern. When an application with dominant communication that is intensive all-to-all submitted to the system, it should be granted with compact allocation with exclusive network provision, so that it won't suffer performance loss due to interference. If job with dominant communication pattern like Nearest Neighbor or others with intensive small scale communication, rather than assign it with unnecessary compact allocation and exclusive network provision, non-contiguous allocation with proper unit size would serve it equally well. No degradation will be introduced to job's performance and system utilization will be improved.

@@

\section{Related Work}
\label{sec:related_work}
%\textbf{P1: tools for monitoring and profiling}\\
There are some tools available for system monitoring and application profiling on HPC systems. Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces. However, recognizing the patterns about application's communication from those traces requires lots of extra effort. Some researchers work on the recognition and characterization of parallel application communication patterns. Oak Ridge National Laboratory has this ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of 12 scientific applications and benchmarks \cite{oxbow}. In their recent work \cite{hpdc2015-ornl}, they demonstrate a new approach to automatically characterizing a parallel application's communication behavior.

@@

%\textbf{P4: Job allocation(little bit)}\\
There are some research work focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs. Hoefler et al. propose to use performance modeling techniques to analysis factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}. However, as the scale of HPC continue grows, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools. Bogdan et al provide a set of guidelines how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}. Dong et al describe IBM Blue Gene/Q's 5D torus interconnect network \cite{Dong-SC11}. They developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and allreduce, to demonstrate the effectiveness of this highly parallel 5D torus network.

@@

\bibitem{hpdc14}
Prisacari, Bogdan and Rodriguez, German and Heidelberger, Philip and Chen, Dong and Minkenberg, Cyriel and Hoefler, Torsten
\newblock Efficient Task Placement and Routing of Nearest Neighbor Exchanges in Dragonfly Networks
\newblock In Proceedings of the 23rd International Symposium on High-performance Parallel and Distributed Computing
\newblock HPDC '14,
\newblock 2014,
\newblock Vancouver, BC, Canada

@@

\bibitem{hpdc2015-ornl}
Philip. C.R, Jeremy S. M, etc
\newblock "Automated Characterization of Parallel Application Communication Patterns"
\newblock HPDC'15

>>>>

Network contention between concurrently running jobs on HPC system is one of the primary causes of performance variability. Optimizing job allocation and avoiding network sharing are proven techniques to alleviate network contention and performance degradation. In this work, we propose a novel job communication pattern aware allocation strategy which serves as a critical module in our future HPC batch scheduler. The novelty of our new strategy is that it makes allocation decisions based on job’s communication pattern in addition to system topology and resource availability information. The scheduler uses job communication patterns information to decide whether its performance relies heavily on the network or not, and what the potential intra- and inter-job interference would be for a given allocation. Thus, the scheduler can pick a preferable resource allocation based on a specific job's communication pattern. We validate the effectiveness of our new design using traces collected from three representative parallel applications. The idea presented in this paper is widely applicable: HPC parallel applications have common communication patterns, and allocating resource with communication pattern awareness can improve performance on these systems with a variety of network topologies.

@@

%\textcolor{red}{
%Partition-based allocation achieves good job performance by sacrificing system utilization, whereas non-contiguous allocation can result in better system utilization but could severely degrade performance of user jobs. What is needed is a more flexible strategy that provides allocations that minimize intra/inter-job contention but allow greater system utilization. In this work, we propose a guideline for future batch scheduling systems, a smart way to allocate resources with the awareness of jobs communication patterns.
%
%Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy. Whether it use partition-based or non-contiguous policy, means it chooses the benefit of either system or job over the other. Some propose compromised policies like buddy system\cite{buddy system paper} still make the allocation for job blindly without knowing job's communication information. 
%
%Although HPC applications complexity continue to grow, people could get more knowledge about the applications communication information with the help from various parallel application profiling tools, such as TAU\cite{tau}, MPIP\cite{mpip} and DUMPI\cite{sst}. Job's communication trace and performance analysis can be get by using such tool at runtime. 
%
%The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions. This is because the communication pattern of parallel applications on HPC system can be summarized into a couple of categories\cite{roth}. The deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern. Job's communication traces study would be extreme helpful to leadership computing facility too. The set of applications running on those leadership computing facilities is usually stable, and the workloads from those HPC systems is usually has great repetitiveness. We analyzed the workload on Mira, an IBM Blue/Gene Q system at Argonne Leadership Computing Facility, found that most jobs require 2k nodes and about 14 particular jobs consist of the majority its workload\ref{fig: repetitiveness of Mira}. In other word, Mira accommodates mostly some particular 14 jobs and most jobs have the same size.
%}

@@

%A parallel application usually conforms to a combination of several basic communication patterns\cite{roth}. At its different execution phases, the application's communication behavior may follow different basic patterns respectively. \textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution. Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc. In this work, we focus on the locality of application's rank-to-rank communication topology graph.

A parallel application usually conforms to a combination of several basic communication patterns \cite{roth}. At its different execution phases, the application's communication behavior may follow different basic patterns respectively. There are many profiling tools \cite{tau}\cite{mpip}\cite{scala}\cite{sst}\cite{oxbow} available to capture information regarding communication patterns of parallel applications. They can help analyze parallel applications' communication behavior, providing information like the percentage of different MPI operations of the applications, communication topology, the amount data transferred between processes.

@@

The second MiniApp we have is CrystalRouter, the extracted communication kernel of the full application Nek5000 \cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory. It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos. CrystalRouter demonstrates the ``Many-to-Many" communication pattern through scalable multi-stage communication process. The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.

The collective communication in CrystalRouter utilizing a recursive doubling approach. Ranks in CrystalRouter conform to a n-dimensional hypercube and recursively splitting into (n-1)-dimensional hypercubes, with communication happening with respect to the links that are severed. The pattern of this recursive communication can be found in Figure \ref{fig: cr communication topology}. By nature of the logarithmic distribution, more data transferring are in the local neighborhood, which consists of 5 ranks.

@@

MultiGrid represents a group of applications whose dominant communication pattern is Many-to-Many (some literatures also refer it as All-to-All \cite{roth}).  Applications with such similar dominant communication pattern are like FillBoundary and MiniDFT \cite{design forward webpage}.

@@

%\textcolor{red}{Apps Summary} These three applications have representative communication patterns that are commonly seen in the HPC system workload. AMG represents a group of applications whose dominant communication pattern is Nearest Neighbor. MultiGrid represents applications whose dominant communication pattern is Many-to-Many (some literatures also refer it to All-to-All\cite{roth}). CrystalRouter represents those applications whose dominant communication is a hybrid of multi-stage local communication and hierarchical global communication.

@@

It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context. One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by user. Even the system administrator is not authorized to make that change. Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference then compare the results with those in the presence of interference. Therefore, we resort to simulation for this work.

@@

A simulation toolkit named CODES enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}. CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}. CODES support both torus and dragonfly network with high fidelity flit-level simulation. CODES has this network workload component that capable of conducting trace-driven simulation. It can take real MPI application trace generated by SST DUMPI \cite{sst} to drive CODES network models. In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems. Since torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
%We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies. The new API can help us to explore the impact of different allocation strategies to those three applications in Section \ref{sec:application study}

@@

The topology of torus network is k-ary n-cube, with $k^n$ nodes in total arranged in an n-dimensional grid having k nodes in each dimension. Each node has 2$\times$n direct linked neighbor nodes. The torus network performance is determined by its dimensionality and link bandwidth. As the increase of the dimensionality of torus network, so does the number of links connected with each node. The increased aggregated bandwidth of each node will definitely reduce the data transfer time of each rank in the application. Figure \ref{fig:dimensionality-study} shows the performance of the AMG, CrystalRouter and MultiGrid on a 2K node torus network model with a 3D torus (16$\times$16$\times$8), a 5D torus (8$\times$4$\times$4$\times$4$\times$4), and a 7D torus (4$\times$4$\times$4$\times$4$\times$2$\times$2$\times$2). The bandwidth of direct link between nodes is 2GiB/s in each direction \cite{bgq}, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus.

We can see from Figure \ref{fig:dimensionality-study} that the data transfer time of those three applications are greatly reduced as the dimensionality increases. The increased aggregated bandwidth of each node can accelerate the data transfer. We allocate each rank within each application linearly on the 3D torus. However, due to the variance of data transfer volume of each rank, AMG's data transfer time has great dispersion, shown as the outliers above the boxes in Figure \ref{fig:dimen-amg}. There is few dispersion of MultiGrid, shown in Figure \ref{fig:dimen-mg}, since data transfer of each rank in MultiGrid varies slightly  between 2.8MB to 2.9MB. CrystalRouter also has few dispersion as shown in Figure \ref{fig:dimen-cr}, this is due to the majority of data transferring is in the local neighborhood, which only take few hops.

@@

Communication variability due to network sharing can cause application performance degradation. \textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application. \textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.

In this section we will study both kinds of interference in the context of torus network. 
Since application's communication pattern won't change with its scale, we choose AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks. And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node (16 $\times$ 16 $\times$ 8) 3D torus network for the experiments.

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/cont/allocation}
        \caption{ }
        \label{fig:cont_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/inter-job/noncont/sharing}
        \caption{ }
        \label{fig:noncont_sub1}
    \end{subfigure}%
    \caption{(a) Illustrates the contiguous allocation for 64-node job in three different shapes. ``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube. (b) Illustrates the non-contiguous allocation. Each job is represented by a specific color. The nodes assigned to different jobs are interleaved, the allocation unit is 4-node.}
\end{figure}

@@

There are lots of research works \cite{jingjin}\cite{ozan} try to come up with better task mapping algorithms to alleviate the intra-job interference, which is out the scope of our work. In this work, we focus on the analysis of such interference with different allocations. To study the intra-job interference of each job, we simulate each job running on 3D torus network with different allocation shapes. We assign each application with three different shapes allocation, i.e. 3D balanced-cube, 3D unbalanced-cube, 2D mesh.

@@

There are lots of research work try to design fancy allocation algorithms to provide application with most compact allocation. However, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application. Figure \ref{fig:shapstudy-amg} shows that for AMG, who has this symmetric local communication pattern through all ranks, spent less time when assigned with 3D-unbalanced allocation than 3D-balanced. And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}. Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message between ranks in MultiGrid. CrystalRouter has with both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.

@@

%We simulate three application running concurrently with different allocation to analysis the interference between them. First, we assign each application with contiguous allocation side by side, and compare their performance with each application running exclusively. For AMG, sharing the network with other application won't affect its performance, since all the data transfer are between neighboring ranks. CrystalRouter and MultiGrid will suffer prolonged communication time when sharing network with others. The results are shown in Figure \ref{fig: concurrent-cont-time-box}.
Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}. Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than partition-based system. Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.

We assign each application with non-contiguous allocation and run them concurrently on the same network. The allocation unit belongs to different jobs are interleaved as shown in Figure \ref{fig:noncont_sub1}. The unit size is critical to the application's performance. Figure \ref{fig: 3apps interjob study} shows the results of each application data transfer time with different allocation unit size.

@@

\subsection{Summary}
\label{sec:summary}

Based on our comprehensive study, we can get observations that 
\begin{enumerate}
    \item The compact allocation may not guarantee the best performance for every application. 
    \item The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefer compact allocation. 
    \item The allocation unit size should be decided according to application's communication pattern. Unit size that big enough to preserve the locality of the application will result better performance. 
\end{enumerate}

\section{Discussion}
\label{sec:discussion}

HPC system usually accommodate many parallel jobs running concurrently. Each job gets the required number of computing nodes by system resource manager. The allocated nodes for each job communicate with each other through system interconnect network in a sharing way, which cause interference as a negative side-effect. This interference can introduce performance degradation to both job and system, such as prolonged job running time and reduced system throughput.

@@

The most promising way to utilize such knowledge is to make smart allocation based on job's dominant communication pattern. Most existing batch schedulers take so few initiative because it fixed-size partition-based or non-contiguous policy, which means it chooses the benefit of either system or job over the other. Some proposed compromised policies \cite{lo}\cite{pascual}\cite{yang-cluster14} to use various allocation unit sizes, but they still  make the allocation for job blindly without knowing job's communication information.

Based on the findings in this work, we propose the idea of smart job allocation with consideration about job's communication pattern. As the dominant communication pattern of parallel applications on HPC system can be summarized into a couple of categories, the deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern. When an application with dominant communication that is intensive all-to-all being submitted to the system, it should be granted with compact allocation with exclusive network provision, so that it won't suffer performance loss due to interference. If job with dominant communication pattern like Nearest Neighbor or others with intensive small scale communication, rather than assign it with unnecessary compact allocation and exclusive network provision, non-contiguous allocation with proper unit size would serve it equally well. No degradation will be introduced to job's performance and system utilization will be improved.

@@

\section{Related Work}
\label{sec:related_work}
%\textbf{P1: tools for monitoring and profiling}\\
There are some tools available for system monitoring and application profiling on HPC systems. Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces. However, recognizing the patterns about application's communication from those traces requires lots of extra effort. Some researchers work on the recognition and characterization of parallel application communication patterns. Oak Ridge National Laboratory has this ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of 12 scientific applications and benchmarks \cite{oxbow}. In their recent work \cite{roth}, they demonstrate a new approach to automatically characterizing a parallel application's communication behavior.

@@

%\textbf{P4: Job allocation(little bit)}\\
There are some research work focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs. Hoefler et al. propose to use performance modeling techniques to analysis factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}. However, as the scale of HPC continue grows, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools. Bogdan et al provide a set of guidelines how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}. Dong et al describe IBM Blue Gene/Q's 5D torus interconnect network \cite{Dong-SC11}. They developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network.

@@

%\bibitem{hpdc14}
%Prisacari, Bogdan and Rodriguez, German and Heidelberger, Philip and Chen, Dong and Minkenberg, Cyriel and Hoefler, Torsten
%\newblock Efficient Task Placement and Routing of Nearest Neighbor Exchanges in Dragonfly Networks
%\newblock In Proceedings of the 23rd International Symposium on High-performance Parallel and Distributed Computing
%\newblock HPDC '14,
%\newblock 2014,
%\newblock Vancouver, BC, Canada

@@
